`r params$project_title`

### 2021-05-06 reports/rmarkdown/01_data_pre_processing.Rmd
Genus_lacto correction

# create genus lacto taxonomy
genus_lacto.fun <- function(df){
  genus_lacto <- df %>%
    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% #ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species),
    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste0(.$Genus_lacto, "crispatus/acidophilus"), 
                                ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
                                       as.character(.$Genus_lacto))) )
  return(genus_lacto)
}

@@

```{r, echo = FALSE}
temp <- map(datasets_, ~.x  %>%
  select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
  filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
  other_taxa.fun(.) %>% # create "other" taxonomy 
  genus_lacto.fun(.)  %>% # create genus lacto taxonomy 
  {. ->> temp_ } %>% 
  aggregate.fun(., "Genus_lacto") 
)

# Extract taxonomic information to file
seq_taxa <- bind_rows(temp_) %>% select(1:11) %>% unique() 
#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_taxa",".csv"),row.names = F )

@@

```{r, echo = FALSE}
PATH <- "../../" #Path to the data
# Load metadata
metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
rownames(metadata) <- as.character(metadata$ID)
# Load datasets
dataset_names <- c("ASV_tissue_B2.csv",
                     "ASV_tissue_B1.csv",
                     "ASV_CVL_V2_B1.csv",
                     "ASV_CVL_V2_B2.csv",
                     "ASV_CVL_V2_C.csv",
                     "ASV_CVL_V3.csv")
#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"data/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  sps <- apply( temp[,6:7] ,1,function(x) paste(x,collapse ="_"))
  not_complete <- grepl("_NA",sps)

  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
  taxa[!not_complete] <- sps[!not_complete]
  taxa <- sub("_"," ",taxa)
  use <- !is.na(taxa) & (taxa != "NA")
  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
  temp <- rowsum(temp[use,],taxa[use])
  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
  return(temp)
})
names(datasets) <- dataset_names
lapply(datasets,dim)
lapply(datasets,colnames)
#arrange all datasets to have the same rownames with all bacteria
common_microbes <- unique(unlist(lapply(datasets,rownames)))
datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)
# datasets <- datasets[3] #in case you wanna test in 1 dataset
# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)
# Extract taxonomic information to file
taxa_info <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"data/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  return(t(temp))
})
taxa_info <- t(as.data.frame(taxa_info))
rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
rownames(taxa_info) <- sub("[.][0123456789]","other ",rownames(taxa_info))
sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
not_complete <- grepl("_NA",sps)
taxa <- apply( taxa_info ,1,function(x) paste(x,collapse ="_"))
taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
taxa[!not_complete] <- sps[!not_complete]
taxa <- sub("_"," ",taxa)
use <- !is.na(taxa) & (taxa != "NA")
taxa[taxa == "NA"] <- "other"
rownames(taxa_info) <- taxa
taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
write.csv2(taxa_info, paste0(PATH,"../results/taxonomy.csv"),row.names = T)
```

>>>>

# create genus lacto taxonomy
genus_lacto.fun <- function(df){
  genus_lacto <- df %>%
    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste(.$Genus_lacto, "crispatus/acidophilus"), 
                                ifelse(grepl("reuteri",.$Species), paste(.$Genus_lacto, "reuteri/oris/frumenti/antri"),
                                  ifelse(grepl("gasseri",.$Species), paste(.$Genus_lacto,"gasseri/johnsonii/taiwanensis"),
                                    ifelse(grepl("murinus",.$Species), paste(.$Genus_lacto,"murinus/animalis/apodemi/salivarius"),
                                      ifelse(grepl("plantarum",.$Species), 
                                           paste(.$Genus_lacto, "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
                                        ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
                                              as.character(.$Genus_lacto))))))) )
  return(genus_lacto)
}

@@

```{r, echo = FALSE}
temp <- map(datasets_, ~.x  %>%
            select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
            filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
            other_taxa.fun(.) %>% # create "other" taxonomy 
            genus_lacto.fun(.)  
            ) %>% # create genus lacto taxonomy 
        {. ->> temp_ } %>% 
        map(., ~aggregate.fun(., "Genus_lacto"))

# Extract taxonomic information to file
seq_taxa <- bind_rows(temp_) %>% select(1:11) %>% unique() 
#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )

@@

```{r, eval=F, echo = FALSE}
PATH <- "../../" #Path to the data
# Load metadata
metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
rownames(metadata) <- as.character(metadata$ID)
# Load datasets
dataset_names <- c("ASV_tissue_B2.csv",
                     "ASV_tissue_B1.csv",
                     "ASV_CVL_V2_B1.csv",
                     "ASV_CVL_V2_B2.csv",
                     "ASV_CVL_V2_C.csv",
                     "ASV_CVL_V3.csv")
#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"data/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  sps <- apply( temp[,6:7] ,1,function(x) paste(x,collapse ="_"))
  not_complete <- grepl("_NA",sps)

  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
  taxa[!not_complete] <- sps[!not_complete]
  taxa <- sub("_"," ",taxa)
  use <- !is.na(taxa) & (taxa != "NA")
  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
  temp <- rowsum(temp[use,],taxa[use])
  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
  return(temp)
})
names(datasets) <- dataset_names
lapply(datasets,dim)
lapply(datasets,colnames)
#arrange all datasets to have the same rownames with all bacteria
common_microbes <- unique(unlist(lapply(datasets,rownames)))
datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)
# datasets <- datasets[3] #in case you wanna test in 1 dataset
# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)
# Extract taxonomic information to file
taxa_info <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"data/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  return(t(temp))
})
taxa_info <- t(as.data.frame(taxa_info))
rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
rownames(taxa_info) <- sub("[.][0123456789]","other ",rownames(taxa_info))
sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
not_complete <- grepl("_NA",sps)
taxa <- apply( taxa_info ,1,function(x) paste(x,collapse ="_"))
taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
taxa[!not_complete] <- sps[!not_complete]
taxa <- sub("_"," ",taxa)
use <- !is.na(taxa) & (taxa != "NA")
taxa[taxa == "NA"] <- "other"
rownames(taxa_info) <- taxa
taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)
```
